---
title: "Matching in R"
author: "Katherine Siegel"
date: "2023-02-16"
output: html_document
---

## Description
Code to demonstrate matching in R. Adapted from the supplementary materials from Butsic, V. et al. (2017): Quasi-experimental methods enable stronger inferences from observational data in ecology. (c) Matthias Baumann (2017-01-10).	

## Set up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

### Load libraries
library(tidyverse)
library(MatchIt)
```

## Simulate data
Simulated data is handy because we know the true effect of the treatment variable.
```{r}
### Function to simulate data and write it as a dataframe
simulate_data <- function(){
  
  ### Create variables within dataframe
  
  ### Make column for observation ID
  df <- data.frame(id = seq(1,1000),
                   
                   ### Add columns for explanatory variables
                   
                   ### Add column for treatment variable
                   fire = c(rep(0,500), rep(1,500)),
                   
                   ### And the rest of the covariates
                   slope = c(runif(500, min = 50, max = 90), 
                             runif(500, min = 65, max = 150)),
                   elevation = c(runif(500, min = 150, max = 185), 
                                 runif(500, min = 165, max = 200)),
                   stream = runif(1000, min = 0, max = 1),
                   
                   ### And the error term
                   error = rnorm(1000, mean = 0, sd = 5))
  
  ### Add a slope*slope variable
  df <- df %>%
    mutate(slope2 = slope^2)
  
  ### Make column for outcome variable (species richness) 
  df <- df %>%
    mutate(species_richness = 1 + 5*fire + 0.07*slope + 0.05*elevation + 2*stream - 0.005*slope2 + error)
  return(df)
}
```
*We know that the true effect of the treatment variable (fire) is a 5x increase in the response variable (species richness).*

## Estimate the effect using ordinary least squares
```{r}
### Write a function to generate data and analyze using OLS 
ols_fun <- function(){
  
  ### Simulate the dataset
  data <- simulate_data()
  
  ### Run OLS regression
  ols <- lm(species_richness ~ fire + slope + elevation + stream, 
            data = data)
  
  ### Extract model coefficients and standard error
  fire_coeff <- coef(summary(ols))["fire", "Estimate"]
  fire_se <- coef(summary(ols))["fire", "Std. Error"]
  list <- list(fire_coeff, fire_se)
}

### Apply the function to 1000 replicates
ols_sim <- replicate(1000, ols_fun())

### Extract the model estimates
ols_fire_est <- unlist(ols_sim[1, ])

### Print mean, standard deviation, minimum, and maximum values for coefficient estimates
c(mean(ols_fire_est), sd(ols_fire_est), 
  min(ols_fire_est), max(ols_fire_est))

### Extract the standard deviations
ols_fire_sd <- unlist(ols_sim[2,])

### Print mean, standard deviation, minimum, and maximum standard deviation of coefficient estimates
c(mean(ols_fire_sd), sd(ols_fire_sd), 
  min(ols_fire_sd), max(ols_fire_sd))
```
*The effect estimated by OLS is incorrect-- it should be 5.*

## Use matching prior to regression
```{r}
### Write a function to generate the data, use matching to subset the data, and run a regression on the matched data
pps_fun <- function(){
  
  ### Simulate the dataset
  data <- simulateData_PPS()
  
  ### Match the data on the observed covariates
  match <- matchit(fire ~ slope + 
                     elevation + 
                     stream, 
                   method = "nearest", 
                   data = data, 
                   distance="probit", 
                   discard="both")
  
  ### Extract the matched data from the full dataset
  matched_data = match.data(match)
  
  ### Run OLS on matched dataset
  ols <- lm(species_richness ~ fire + slope + 
              elevation + stream, 
            data = matched_data)
  
  ### Extract model coefficients
  fire_coeff <- coef(summary(ols))["fire", "Estimate"]
  fire_se <- coef(summary(ols))["fire", "Std. Error"]
  list <- list(fire_coeff, fire_se) 
}
pps_sim <- replicate(1000, pps_fun())
pps_fire_est <- unlist(pps_sim[1,])
c(mean(pps_fire_est), sd(pps_fire_est), min(pps_fire_est), max(pps_fire_est))
pps_fire_sd <- unlist(pps_sim[2,])
c(mean(pps_fire_sd), sd(pps_fire_sd), min(pps_fire_sd), max(pps_fire_sd))
```

