---
title: "Regression discontinuity design in R"
output: html_document
date: "2023-03-20"
---

## Description
Code to demonstrate matching in R. Adapted from the supplementary materials from Butsic, V. et al. (2017): Quasi-experimental methods enable stronger inferences from observational data in ecology. (c) Matthias Baumann (2017-01-10).	

In the Butsic et al. paper, they used the example of the impact of wildfire on species richness. Here, we will simulate data with a known treatment effect of fire on species richness. We will then compare the estimated effect we get through a naive ordinary least squares (OLS) regression approach to the effect we estimate when we use RDD methods .

## Set up
Load required packages. In this demo, we will use the package "rdrobust" for the matching process.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

### Install packages if you need to
install.packages("rdrobust",
                 repos = "http://cran.us.r-project.org")
install.packages("rddensity",
                 repos = "http://cran.us.r-project.org")

### Load libraries
library(tidyverse)
library(rdrobust)
library(rddensity)
```

## Simulate data
Simulated data is handy because we know the true effect of the treatment variable. Here, we'll write a function to simulate a dataset where we know the true effect of fire on species richness.
```{r}
### Function to simulate data and write it as a dataframe
simulate_data <- function(){
  
  ### Create variables in a dataframe
  
  ### Make column for observation ID
  df <- data.frame(id = seq(1, 1000),
                   
                   ### Add columns for explanatory variables
                   
                   ### Add column for treatment variable
                   fire = c(rep(0, 500), rep(1, 500)),
                   
                   ### And the rest of the covariates
                   distance = c(runif(500, min = -100, max = 0), 
                             runif(500, min = 0, max = 100)),
                   slope = runif(1000, min = 50, max = 100),
                   elevation = runif(1000, min = 150, max = 185),
                   stream = runif(1000, min = 0, max = 1),
                   
                   ### And the error term
                   error = rnorm(1000, mean = 0, sd = 2))
  
  ### Add land use history variable, depends on outcomes of distance
  df <- df %>%
    mutate(land_use = ifelse(distance > -50 & distance < 5, 
                             1, 
                             0))
  
  
  
  ### Make column for outcome variable (species richness) 
  df <- df %>%
    mutate(species_richness = 1 + 5*fire + 0.7*slope + 0.05*elevation + 2*stream + 5*land_use + error)
  return(df)
}
```
*We know that the true effect of the treatment variable (fire) is a 5x increase in the response variable (species richness).*

## Estimate the effect using ordinary least squares
```{r}
### Write a function to generate data and analyze using OLS 
ols_fun <- function(){
  
  ### Simulate the dataset
  data <- simulate_data()
  
  ### Run OLS regression
  ols <- lm(species_richness ~ fire + slope + elevation + stream, 
            data = data)
  
  ### Extract model coefficients and standard error
  fire_coeff <- coef(summary(ols))["fire", "Estimate"]
  fire_se <- coef(summary(ols))["fire", "Std. Error"]
  list <- list(fire_coeff, fire_se)
}

### Apply the function to 100 replicates
ols_sim <- replicate(100, ols_fun())

### Extract the model estimates
ols_fire_est <- unlist(ols_sim[1, ])

### Print mean, standard deviation, minimum, and maximum values for coefficient estimates
c(mean(ols_fire_est), sd(ols_fire_est), 
  min(ols_fire_est), max(ols_fire_est))

### Extract the standard deviations
ols_fire_sd <- unlist(ols_sim[2, ])

### Print mean, standard deviation, minimum, and maximum standard deviation of coefficient estimates
c(mean(ols_fire_sd), sd(ols_fire_sd), 
  min(ols_fire_sd), max(ols_fire_sd))
```
*The effect estimated by OLS is incorrect-- it should be 5.*

## Use regression discontinuity
```{r}
### Simulate the dataset
data_rdd <- simulate_data()

### look at the distribution of sample points along the running variable
# ggplot(data = data_rdd, aes(x = distance)) + 
#   geom_histogram() +
#   geom_vline(xintercept = 0)
# 
# ### look at the distribution of the outcome variable along the running variable
# ggplot(data = data_rdd, aes(x = distance, y = species_richness)) +
#   geom_point() +
#   geom_vline(xintercept = 0)


  

```
